---
title: "disc_week_6"
author: "Hanying Jiang"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css", "cols.css"]
    lib_dir: libs
    nature:
      beforeInit: "cols_macro.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
    seal: false  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

# Week 6
## Today's plan
 - Solve question 1 by groups (15 - 20 min).
 - My explanation
 - Solve question 1 by groups. (10 - 15 min).
 - My explanation
 - A quick review

## Main topics
 - Linear Contrast.
 
---
### 1
Consider a completely randomized design with five treatments, four units per treatment, and treatment means

|   treatment |  sample mean |
| :---------: | :----------: |
|           1 |          3.1 |
|           2 |          9.8 |
|           3 |          8.1 |
|           4 |          8.2 |
|           5 |          7.6 |


The MSE is 4.0.
```{r}
sampleMeans <- c(3.1, 9.8, 8.1, 8.2, 7.6)
MSE <- 4.0
```

(a) Construct an ANOVA table for this experiment and test the null hypothesis that all treatments have the same mean.

(b) Test the null hypothesis that the average response in treatments 1 and 2 is the same as the average response in treatments 3, 4, and 5. Use $\alpha = 0.05$.

(c) Use the HSD procedure to compare the means of the five treatments. Use $\alpha = 0.05$.
---

Consider a completely randomized design with five treatments, four units per treatment

```{r}
sampleMeans <- c(3.1, 9.8, 8.1, 8.2, 7.6)
MSE <- 4.0
```

(a) Construct an ANOVA table for this experiment and test the null hypothesis that all treatments have the same mean.

```{r, include=T, echo=F}
source = c('Treatment', 'error', 'total')
df = c('?', '?', '?')
ss = c('?', '?', '?')
ms = c('?', '4.0', '')
f = c('?', '', '')
p = c('?', '', '')

kable(noquote(cbind(source, ss, df,  ms, f, p)))
```

---

Consider a completely randomized design with five treatments, four units per treatment

```{r}
sampleMeans <- c(3.1, 9.8, 8.1, 8.2, 7.6)
MSE <- 4.0
```

(a) Construct an ANOVA table for this experiment and test the null hypothesis that all treatments have the same mean.

$$
a = 5, n = 4, N = a*n=20
$$
$$
df \; total = N-1=19 
$$
$$
df \; Treatment = a-1=4
$$

$$
df \; Error = N-a=15
$$

$$
SSE = MSE \times df \; Error = 4.0*15=60
$$

---

Consider a completely randomized design with five treatments, four units per treatment

```{r}
sampleMeans <- c(3.1, 9.8, 8.1, 8.2, 7.6)
MSE <- 4.0
```

(a) Construct an ANOVA table for this experiment and test the null hypothesis that all treatments have the same mean.

```{r}
n <- 4 # observations per treatment
a <- length(sampleMeans)  # 5 recipes
N <- n*a  # total sample size

dfE <- N - a
dfTr <- a - 1
dfT <- N - 1

SSE <- dfE * MSE

c(dfE, dfTr, dfT, SSE)
```


```{r, include=T, echo=F}
source = c('Treatment', 'error', 'total')
df = c(4, 15, 19)
ss = c('?', '60', '?')
ms = c('?', '4.0', '')
f = c('?', '', '')
p = c('?', '', '')

kable(noquote(cbind(source, ss, df,  ms, f, p)))
```

---

Consider a completely randomized design with five treatments, four units per treatment

```{r}
sampleMeans <- c(3.1, 9.8, 8.1, 8.2, 7.6)
MSE <- 4.0
```

(a) Construct an ANOVA table for this experiment and test the null hypothesis that all treatments have the same mean.

$$
SS_{Treatment} = n \sum{(\bar{y}_i - \bar{y})^2}
$$
```{r}
SSTr <- n*sum((sampleMeans - mean(sampleMeans))^2)

MSTr <- SSTr / dfTr

SST <- SSTr + SSE
MST <- SST / dfT

c(SSTr, MSTr, SST, MST)
```

```{r, include=T, echo=F}
source = c('Treatment', 'error', 'total')
df = c(4, 15, 19)
ss = c(SSTr, SSE, SST)
ms = c(MSTr, MSE, MST)
f = c('?', '', '')
p = c('?', '', '')

kable(data.frame(source, ss, df,  ms, f, p))
```

---

Consider a completely randomized design with five treatments, four units per treatment

```{r}
sampleMeans <- c(3.1, 9.8, 8.1, 8.2, 7.6)
MSE <- 4.0
```

(a) Construct an ANOVA table for this experiment and test the null hypothesis that all treatments have the same mean.

```{r}
f <- MSTr / MSE
p <- pf(f, df1 = a-1, df2 = N-a, lower.tail = F) 
p <- round(p, 5)
f
p
```

```{r, include=T, echo=F}
source = c('Treatment', 'error', 'total')
df = c(4, 15, 19)
ss = c(SSTr, SSE, SST)
ms = c(MSTr, MSE, MST)
f_ = c(as.character(f), '', '')
p_ = c(as.character(p), '', '')
aovtable = data.frame(source, ss, df,  ms, f_, p_)
kable(data.frame(source, ss, df,  ms, f_, p_))
```

---

Consider a completely randomized design with five treatments, four units per treatment

```{r, include=T, echo=F}
kable(aovtable)
```

(b) Test the null hypothesis that the average response in treatments 1 and 2 is the same as the average response in treatments 3, 4, and 5.

---

Consider a completely randomized design with five treatments, four units per treatment

```{r, include=T, echo=F}
kable(aovtable)
```

(b) Test the null hypothesis that the average response in treatments 1 and 2 is the same as the average response in treatments 3, 4, and 5.

$$
H_0: \frac{\mu_1+\mu_2}{2} = \frac{\mu_3+\mu_4+\mu_5}{3} 
$$

Or 

$$
H_0: {1 \over  2}\mu_1 + {1 \over  2}\mu_2 - {1 \over  3}\mu_3- {1 \over  3}\mu_4- {1 \over  3}\mu_5 = 0
$$

$$
H_A: {1 \over  2}\mu_1 + {1 \over  2}\mu_2 - {1 \over  3}\mu_3- {1 \over  3}\mu_4- {1 \over  3}\mu_5 \neq 0
$$

Notice

$$
{1 \over  2} + {1 \over  2} - {1 \over  3}- {1 \over  3}- {1 \over  3}=0
$$

---
(b) Test the null hypothesis that the average response in treatments 1 and 2 is the same as the average response in treatments 3, 4, and 5.

$$
t = \frac{\sum{c_i \bar{y}_i} - 0}{\sqrt{\frac{MSE}{n}\sum{c_i^2}}}
$$

```{r}
c1 <- c(1/2, 1/2, -1/3, -1/3, -1/3) # contrast coefficient
C1 <- sampleMeans %*% c1 # sum(c1_i * y_i bar)

seC1 <- sqrt(MSE/n * sum(c1^2))

t1 <- C1 / seC1

pt(abs(t1), df = N-a, lower.tail = F) * 2
```

```{r}
pt(abs(t1), df = N-a, lower.tail = F) * 2 < 0.05
```


---

(c) Use the HSD procedure to compare the means of the five treatments. Use $\alpha = 0.05$.

$$
HSD = q_{\alpha}(a,f) \sqrt{\frac{MS_E}{n}}
$$

```{r}
q <- qtukey(0.05, nmeans = a, df = N-a, lower.tail = F)
HSD <- q * sqrt(MSE/n)
HSD
```

| treatment |sample mean |Gap  |
|:---------:|:----------:|:---:|
|         1 |        3.1 |     |
|         5 |        7.6 | 4.5 |
|         3 |        8.1 | 0.5 |
|         4 |        8.2 | 0.1 |
|         2 |        9.8 | 1.6 |
|           |            |     |


---

### Question 2

The data below give the mass of rotten tomato 4 days after a tomato was subjected to injection of 4 different bacteria treatments.  Treatment 1 is a control (no bacteria injected), Treatment 2 is bacteria type A injected, Treatment 3 is bacteria type B injected, Treatment 4 is both bacteria type A and B injected.

```{r}
mass <- c(18.9, 18.8, 19.5, 21.3, 18.7, 19.2, 19.9, 21.1)
treatment <- factor(rep(1:4, times = 2))
dat <- data.frame(treatment, mass)
head(dat)
```

Test the hypotheses corresponding to the following contrasts with the following coefficients:

- C1: 1, 1, -1, -1
- C2: 1, -1, 1, -1
- C3: 1, -1, -1, 1

Interpret the results in the context of the problem.

---

First, define the contrast:

```{r}
contrasts <- matrix(c(1, 1, -1, -1, 
                      1, -1, 1, -1,
                      1, -1, -1, 1), byrow = T, nrow =3)
contrasts
```
Notice that every row is a contrast. byrow determins the order we fill the number into the matrix. nrow determines the shape of the matrix.

---

```{r}
library(gmodels)
fit <- aov(mass ~ treatment, data = dat)

# fit.contrast(model, varname, coefficients, confidence level)
fit.contrast(fit, "treatment", contrasts, conf.int = 0.95)

```

---

| treatment | bacteria A | bacteria B |
|:---------:|:----------:|:----------:|
|    1      |     No     |     No     |
|    2      |    Yes     |     No     |
|    3      |     No     |    Yes     |
|    4      |    Yes     |    Yes     |

C1:  

Coefficients: 1, 1, -1, -1

Hypothesis tested:

$$
H_0: \mu_1 + \mu_2 = \mu_3 + \mu_4
$$

Interpretation: C1 is a contrast that tests whether there is a difference in the average rot of tomatoes treated with bacteria B compared to those not treated with bacteria B.

p-value = 0.0006

Conclusion: we have strong evidence that there existes difference in the average rot of tomatoes treated with bacteria B compared to those not treated with bacteria B.

---

| treatment | bacteria A | bacteria B |
|:---------:|:----------:|:----------:|
|    1      |     No     |     No     |
|    2      |    Yes     |     No     |
|    3      |     No     |    Yes     |
|    4      |    Yes     |    Yes     |

C2:  

Coefficients: 1, -1, 1, -1

Hypothesis tested:

$$
H_0: \mu_1 + \mu_3 = \mu_2 + \mu_4
$$

Interpretation: C2 is a contrast that tests whether there is a difference in the average rot of tomatoes treated with bacteria A compared to those not treated with bacteria A.

p-value = 0.0060

Conclusion: we have strong evidence that there existes difference in the average rot of tomatoes treated with bacteria A compared to those not treated with bacteria A.

---


| treatment | bacteria A | bacteria B |
|:---------:|:----------:|:----------:|
|    1      |     No     |     No     |
|    2      |    Yes     |     No     |
|    3      |     No     |    Yes     |
|    4      |    Yes     |    Yes     |

C3:  

Coefficients: 1, -1, -1, 1

Hypothesis tested:

$$
H_0: \mu_1 - \mu_2 = \mu_3 - \mu_4
$$
and also can be written:
$$
H_0: \mu_4 - \mu_2 = \mu_3 - \mu_1
$$
Interpretation: C3 is a contrast that tests whether the effect of bacteria B depends on whether bacteria A is also present.  Equivalently, it tests whether the effect of bacteria A depends on whether bacteria B is also present.


p-value = 0.014

Conclusion: we have evidence that the effect of bacteria B depends on whether bacteria A is also present. Equivalently, it tests whether the effect of bacteria A depends on whether bacteria B is also present.

---

# quick review

What is a contrast?

$$
\sum_ i ^ a {c_i \mu_i}, \, where \, \sum_i^a {c_i} = 0.
$$
---

T test

$$
H_0: \sum_i^a c_i\mu_i = \mu_c^0  \text{ vs. } H_A: \sum_i^a c_i\mu_i \neq \mu_c^0
$$

Test statistic:

$$
t_0 = \frac{\sum{c_i \bar{y}_i} - \mu_c^0}{\sqrt{\frac{MSE}{n} \sum{c_i^2}}}
$$

If $H_0$ is true, $t_0 \sim t_{N-a}$.

How to perform in R:

```{r, eval = FALSE}
# Assume your data is saved in a data frame named as dat
# dat has two columns: treatment stores the treatment types and measurements stores the outcomes
# Make sure treatment is factor!!! If it's not, try:
# dat$treatment = factor(dat$treatment )

# If you only test one contrast, variable contrasts can be a vector which stores the coefficients
# If you are testing multiple contrasts at the same time, construct a variable contrasts, which is a matrix whose each row is the coefficients for one constrast

library(gmodels)
fit <- aov(measurements ~ treatment, data = dat)
fit.contrast(fit, "treatment", contrasts, conf.int = 0.95)
```


---

F test

$$
H_0: \sum_i^a c_i\mu_i = 0  \text{ vs. } H_A: \sum_i^a c_i\mu_i \neq 0
$$
Test statistic:

$$
F = \frac{SS_C/1}{MSE}= \frac{(\sum{c_i \bar{y}_i})^2}{\frac{1}{n} \sum{c_i^2}MSE}
$$

Compare with t:

$$
t_0 = \frac{\sum{c_i \bar{y}_i}}{\sqrt{\frac{MSE}{n} \sum{c_i^2}}}
$$

We can see $F=t_0^2$. And these two tests will give the same conclusion.

---

Pay attention to the computation.



