---

title: disc_week_12
author: Hanying Jiang
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css", "cols.css"]
    lib_dir: libs
    nature:
      titleSlideClass: ["right", "top", "my-title"]
      beforeInit: "cols_macro.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
     
---

#### 6.24

An experiment was run in a semiconductor fabrication plant in an effort to increase yield.  Five factors, each at two levels, were studied.  The factors (and levels) were A = aperture setting (small, large), B = exposure time (20% below nominal, 20% above nominal), C = development time (30 s, 45 s), D = mask dimension (small, large), and E = etch time (14.5 min, 15.5 min).  The unreplicated 25 design was run. Results are saved in data1.csv.

(a)	Construct a normal probability plot of the effect estimates.  Which effects appear to be large?

(b)	Fit the model that includes only the terms you think are different than zero from (a). Conduct an analysis of variance to confirm your findings for part (a).

(c)	Write down the regression model relating yield to the significant process variables.

(d)	Plot the residuals on normal probability paper.  Is the plot satisfactory?

(e)	Plot the residuals versus the predicted yields and versus each of the five factors. Comment on the plots.

(f)	Interpret any significant interactions.

(g)	What are your recommendations regarding process operating conditions?

---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
data1 <- read.csv("data1.csv")
head(data1)
```

---

Encode the factor level by -1 and 1:

```{r}
library(tidyverse)
data1 <- read.csv("data1.csv")
data1 <- data1 %>% mutate(
  A = ifelse(Aperture == "small", -1, 1),
  B = ifelse(Exposure.Time == -20, -1, 1),
  C = ifelse(Develop.Time == 30, -1, 1),
  D = ifelse(Mask.Dimension == "Small", -1, 1),
  E = ifelse(Etch.Time == 14.5, -1, 1)
)
head(data1)
```


---

```{r}
m <- lm(Yield ~ A * B * C * D * E, data = data1)
effects <- 2 * coef(m)[-1]

daniel_plot <- function(effects, p = c(0.3, 0.7)) { 
  qq <- qqnorm(effects, datax = TRUE)
  qqline(effects, col = "red", probs = p, datax = TRUE)
  text(qq$x, qq$y, names(effects), pos=1)
}
daniel_plot(effects)
```

From the normal probability plot of effects shown below, effects A, B, C, and the AB interaction appear to be large.

---

(b)	Fit the model that includes only the terms you think are different than zero from (a). Conduct an analysis of variance to confirm your findings for part (a).

```{r}
m2 <- lm(Yield ~ Aperture + Exposure.Time + Develop.Time + Aperture*Exposure.Time, data = data1)
anova(m2)
```

Effects A, B, C, and the AB interaction are significant, which agrees with the conclusion in (a).

---

(c)	Write down the regression model relating yield to the significant process variables.

```{r}
coef(m2)
```

The model is
$$
\begin{aligned}
Yield = &12.2187500 - 11.8125000*I(Aperture = small) + 1.0468750 *Exposure.Time \\
&-0.3968750 *I(Aperture = small) *Exposure.Time + 0.6458333*Develop.Time
\end{aligned}
$$

When aperture is small, the model can be simplified as:
$$
\begin{aligned}
 Yield &= 12.2187500  -11.8125000 + 1.0468750 *Exposure.Time -0.3968750 *Exposure.Time +  0.6458333*Develop.Time\\
 &=0.40625+0.6500*Exposure.Time+0.6458333*Develop.Time
\end{aligned}
$$

When aperture is large, the model can be simplified as:

$$
Yield = 12.2187500  +1.0468750 *Exposure.Time  +  0.6458333*Develop.Time
$$


---

(d)	Plot the residuals on normal probability paper.  Is the plot satisfactory?

```{r, echo = FALSE}
plot(m2, which = 2)
```

There is nothing unusual about this plot.


---

(e)	Plot the residuals versus the predicted yields and versus each of the five factors. Comment on the plots.

The plot of residual versus exposure time shows some very slight inequality of variance.  There is no strong evidence of a potential problem.

```{r, echo = FALSE}
data1$residuals <- resid(m2)
for (factor in c( "Exposure.Time")){
  plot_title <- sprintf("Residuals vs. %s", factor)
  fig <- data1 %>% ggplot(aes_string(x = factor, y = "residuals")) + geom_point() + ggtitle(plot_title)
  
  print(fig)
}

```


---

(f)	Interpret any significant interactions.

```{r, echo = FALSE, out.height=500}
interaction.plot(x.factor = data1$Exposure.Time,
                 trace.factor = data1$Aperture,
                 response = data1$Yield,
                 fun = mean, 
                 ylab = "Yield",
                 xlab = "Exposure Time",
                 col = c("red", "blue"),
                 trace.label = "Aperture")
```

Factor A (Aperture) does not have as large an effect when B (Exposure.Time) is at its low level (-20) as it does when B is at its high level (20).

---

(g)	What are your recommendations regarding process operating conditions?


To achieve the highest yield:

 - Effect of B is positive: run B at the high level.
 - For AC, from interaction plot, we should both run at high level.

---

#### 7.3

Consider the data from the first replicate of Problem 6.5.  Suppose that these observations could not all be run using the same bar stock.  Set up a design to run these observations in two blocks of four observations each with ABC confounded.  Analyze the data.

```{r}
library(knitr)
data2 <-data.frame(
  A = rep(c(-1,1),times = 4),
  B = rep(rep(c(-1,1),each = 2), times = 2),
  C = rep(c(-1,1), each = 4),
  label = c("1","a","b","ab","c","ac","bc","abc"),
  result = c(22,32,35,55,44,40,60,39)
)
kable(data2)
```

---

Design:
```{r, echo = FALSE}
block_design <- data.frame(block1 = c("1","ab","ac","bc"), block2 = c("a","b","c","abc"))
kable(block_design)
```

---

```{r, echo=FALSE}
daniel_plot <- function(effects, p = c(0.3, 0.7)) { 
  qq <- qqnorm(effects, datax = TRUE)
  qqline(effects, col = "red", probs = p, datax = TRUE)
  text(qq$x, qq$y, names(effects), pos=1)
}
```

```{r, out.height=390}
data2$block <- c(1, 2, 2, 1, 2, 1, 1, 2)
m3 <- lm(result ~ A * B * C + block, data = data2)

effects <- 2 * coef(m3)[-1]
daniel_plot(effects)
```

From the normal probability plot of effects, B, C, and the AC interaction are significant.  Factor A was included in the analysis of variance to preserve hierarchy.

---

```{r}
m4 <- lm(result ~ A * C + B+ block, data = data2)
anova(m4)
```

This design identifies the same significant factors as Problem 6.5.



