---
title: "disc_week_5"
author: "Hanying Jiang"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css", "cols.css"]
    lib_dir: libs
    nature:
      beforeInit: "cols_macro.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
    seal: false  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
library(knitr)
library(kableExtra)
```
```{r,message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(agricolae)
```
# Week 5

## Today's plan
 - Try to solve question 3.9 and 3.10 by groups (20 - 25 min).
 - I'll show you how to solve it.
 - A quick review

## Main topics
 - How to implement anova and check its assumptions in R. 
 - How to use Fisher's LSD to compare pairs of treatments.
 - How to use Tukey's HSD to compare pairs of treatments.
---

3.9

The tensile strength of Portland cement is being studied.  Four different mixing techniques can be used economically.  A completely randomized experiment was conducted and the following data were collected.

```{r, echo = FALSE}
dat = data.frame(Mixing_Technique = c(1,2,3,4),
                 t1 = c(3129, 3200, 2800, 2600),
                 t2 = c(3000, 3300, 2900, 2700),
                 t3 = c(2865, 2975, 2985, 2600),
                 t4 = c(2890, 3150, 3050, 2765))
dat %>% kable(col.names = NULL) %>%add_header_above(c("Mixing Technique" = 1, "Tensile strength (lb/in^2)" = 4))
```
(a) Test the hypothesis that mixing techniques affect the strength of the cement.  Use $\alpha = 0.05$.


(c) Use the Fisher LSD method with α=0.05 to make comparisons between pairs of means.



(d) Construct a QQ plot of the residuals.  What conclusion would you draw about the validity of the normality assumption?  


(e)	Plot the residuals versus the predicted tensile strength.  Comment on the plot.


(f)	Prepare a scatter plot of the results to aid the interpretation of the results of this experiment.

(No need to solve (b). And (d) requires constructing QQ plot, which is slightly different from the textbook.)



3.10 (a)	Rework part (c) of Problem 3.9 using Tukey’s test with α = 0.05.  Do you get the same conclusions from Tukey’s test that you did from the graphical procedure and/or the Fisher LSD method?

3.10 (b)	Explain the difference between the Tukey and Fisher procedures.
---
### (a) Test the hypothesis that mixing techniques affect the strength of the cement.  Use $\alpha = 0.05$.

$$
H_0: \mu_1 = \mu_2 = \mu_3 = \mu_4\\
$$

$$
H_A: \mu_i \neq \mu_j \text{ for at least one pair (i, j)}
$$
---

### (a) Test the hypothesis that mixing techniques affect the strength of the cement.  Use $\alpha = 0.05$.

```{r}
dat = data.frame(technique = factor(rep(1:4, 4)), strength = c(3129, 3200, 2800, 2600, 3000, 3300, 2900, 2700, 2865, 2975, 2985, 2600, 2890, 3150, 3050, 2765))
dat
```

---
### (a) Test the hypothesis that mixing techniques affect the strength of the cement.  Use $\alpha = 0.05$.

```{r}
dat = data.frame(technique = factor(rep(1:4, 4)), strength = c(3129, 3200, 2800, 2600, 3000, 3300, 2900, 2700, 2865, 2975, 2985, 2600, 2890, 3150, 3050, 2765))
m <- aov(strength ~ technique, data = dat)
anova(m) 
```

---

### (c) Use the Fisher LSD method with α=0.05 to make comparisons between pairs of means.

$$
LSD = t_{\alpha/2,N-a}\sqrt{\frac{2MS_E}{n }}
$$
---

(c) Use the Fisher LSD method with α=0.05 to make comparisons between pairs of means.

```{r}
out <- LSD.test(m, "technique",  alpha = 0.05, p.adj="none")
# m: model
# "technique": treatment
out
```

---
(c) Use the Fisher LSD method with α=0.05 to make comparisons between pairs of means.

```{r}
out <- LSD.test(m, "technique",  alpha = 0.05, p.adj="none")
# m: model
# "technique": treatment
LSD = out$statistics$LSD
print(paste("LSD = ", LSD,sep = ""))
```

---
(c) Use the Fisher LSD method with α=0.05 to make comparisons between pairs of means.

```{r}
out$groups
```

Compare the differences between groups with LSD.

```{r,echo = FALSE}
group_mean = out$groups$strength[order(rownames(out$groups))]
for (i in 1:3){
  for(j in (i+1):4){
    diff = abs(group_mean[j] - group_mean[i])
    larger = ifelse(diff>LSD,">","<")
    comparison = sprintf("Treatment %s vs. Treatment %s = |%s - %s| = %s %s %s",j,i, group_mean[j],group_mean[i],diff,larger,LSD)
    print(comparison)
    
  }
}
```

---
(d) Construct a QQ plot of the residuals.  What conclusion would you draw about the validity of the normality assumption?  

```{r}
qqnorm(resid(m))
qqline(resid(m), col = "red")
```

---
## (e)	Plot the residuals versus the predicted tensile strength.  Comment on the plot.
Question:  
 - What assumption is checked by this plot?
 - What else plot is also used to check this assumption?

---

(e)	Plot the residuals versus the predicted tensile strength.  Comment on the plot.

```{r}
dat$predicted = predict(m)
dat$residual = resid(m)
ggplot(dat) + geom_point(aes(x = predicted, y = residual))
```

---

(f)	Prepare a scatter plot of the results to aid the interpretation of the results of this experiment.

```{r,echo = FALSE, message=FALSE}
mean_strength = out$means %>% mutate(technique = 1:4)
dat %>% ggplot(aes(x = technique, y = strength)) +
  geom_point() + 
  geom_segment(aes(x = 1, y = out$means$LCL[1], xend = 1, yend = out$means$UCL[1]), color="red") +
  geom_segment(aes(x = 2, y = out$means$LCL[2], xend = 2, yend = out$means$UCL[2]), color="red") + 
  geom_segment(aes(x = 3, y = out$means$LCL[3], xend = 3, yend = out$means$UCL[3]), color="red") + 
  geom_segment(aes(x = 4, y = out$means$LCL[4], xend = 4, yend = out$means$UCL[4]), color="red") + 
  geom_line(data=mean_strength,aes(x = technique, y = strength),linetype = "dotted") + geom_point()
```

---

### 3.10(a)	Rework part (c) of Problem 3.9 using Tukey’s test with α = 0.05.  Do you get the same conclusions from Tukey’s test that you did from the graphical procedure and/or the Fisher LSD method?

$$
HSD = q_{\alpha}(a,f) \sqrt{\frac{MS_E}{n}}
$$
```{r}
qtukey(0.05, nmeans = 4, df = 12, lower.tail = F)*sqrt(12826/4)
# a = 4
# f = df ERROR = N - a = 16 - 4 = 12
# MSE = 12826
```

```{r}
(HSD.test(m, "technique", alpha = 0.05))$statistics$MSD
```




---

### 3.10(a)	Rework part (c) of Problem 3.9 using Tukey’s test with α = 0.05.  Do you get the same conclusions from Tukey’s test that you did from the graphical procedure and/or the Fisher LSD method?

```{r}
TukeyHSD(m)
```

---

```{r}
TukeyHSD(m)
```

Fisher LSD method

```{r,echo = FALSE}
group_mean = out$groups$strength[order(rownames(out$groups))]
for (i in 1:3){
  for(j in (i+1):4){
    diff = abs(group_mean[j] - group_mean[i])
    larger = ifelse(diff>LSD,">","<")
    comparison = sprintf("Treatment %s vs. Treatment %s = |%s - %s| = %s %s %s",j,i, group_mean[j],group_mean[i],diff,larger,LSD)
    print(comparison)
    
  }
}
```

---
### Comparison of the conclusion

```{r,echo = FALSE}
gro = c()
for (i in 1:3){
  for(j in (i+1):4){
    gro = c(gro,sprintf("%s vs. %s",i,j))
  }
}
sig = c("different", "not different")
fis = sapply(c(1,2,1,1,1,1), FUN = function(x) sig[x])
tuk = sapply(c(2,2,1,2,1,1), FUN = function(x) sig[x])
tem = c("same conclusion", "different conclusion")
comp = sapply(c(2,1,1,2,1,1), FUN = function(x) tem[x])
compari = data.frame(groups = gro,Fisher = fis, Tukey = tuk, comparison = comp )
compari %>% kable()
```

---

## (b)	Explain the difference between the Tukey and Fisher procedures.

$$
LSD = t_{\alpha/2,N-a}\sqrt{\frac{2MS_E}{n }}
$$

$$
HSD = q_{\alpha}(a,f) \sqrt{\frac{MS_E}{n}}
$$

Both Tukey and Fisher utilize a single critical value; however, Tukey’s is based on the studentized range statistic while Fisher’s is based on $t$ distribution.

Also, Fisher's is based on the distribution of 
$$
\frac{\bar{y}_i - \bar{y}_j} {\sqrt{2 MS_E/n}}.
$$




Tukey's is based on the distribution of 

$$
\frac{\bar{y}_{max}-\bar{y}_{min}}{\sqrt{MS_E/n}}.
$$

---

# Review

dat is a two-column data frame. technique is the factor, strength is the observation.

```{r}
dat = data.frame(technique = factor(rep(1:4, 4)), strength = c(3129, 3200, 2800, 2600, 3000, 3300, 2900, 2700, 2865, 2975, 2985, 2600, 2890, 3150, 3050, 2765)) 
dat
```


---

### scatter plot for explorary analysis:

```{r, fig.width = 5, fig.height = 4}
ggplot(dat) + geom_point(aes(x = technique, y = strength) )
```

---

### ANOVA:

```{r}
# model
m <- aov(strength ~ technique, data = dat)
anova(m) 
```

---

### Check assumption: normality of errors


```{r, fig.width = 5, fig.height = 4}
# QQ plot of residuals
qqnorm(resid(m))
qqline(resid(m), col = "red")
```

What we expect: a straight line.

---

### Check assumption: equal variance

```{r,eval = FALSE}
dat$predicted = predict(m)
dat$residual = resid(m)
# residual vs. predicted value
ggplot(dat) + geom_point(aes(x = predicted, y = residual))
#residuals vs. explanatory variables (factor)
ggplot(dat) + geom_point(aes(x = technique, y = residual))
```

.pull-left[
```{r, fig.width = 5, fig.height = 4, echo=F}
dat$predicted = predict(m)
dat$residual = resid(m)
# residual vs. predicted value
ggplot(dat) + geom_point(aes(x = predicted, y = residual)) +
  labs(title = "Residual against fit")
```
]
.pull-right[
```{r, fig.width = 5, fig.height = 4, echo=F}
ggplot(dat) + geom_point(aes(x = technique, y = residual)) +
  labs(title = "Residual against factor")
```
]


What we expect: no obvious trend.

---

### compare pairs of treatments.

Fisher's LSD:

```{r}
out <- LSD.test(m, "technique",  alpha = 0.05, p.adj="none")
LSD = out$statistics$LSD
LSD
```

Then: compare with differences between group means. Group means can be found by:

```{r}
out$groups
```

---

Tukey's HSD:

```{r}
out2 <- HSD.test(m, "technique", alpha = 0.05)
HSD = out2$statistics$MSD
HSD
```

Then we can compare it with differences between group means. But we could also use:

```{r}
TukeyHSD(m)
```





